---
author:
  - Sylvia D. Elvira
  - Aly Lamuri
  - Khamelia Malik
  - Murdani Abdullah
  - Petrin R. Lukman
  - Hamzah Shatri
title: Psychological Distress among Greater Jakarta Area Residents
subtitle: COVID-19 Pandemic and Community Containment
output:
  ioslides_presentation:
    self_contained: false
    widescreen: true
---

```{r init, echo=FALSE}
options(digits=2)
options("kableExtra.html.bsTable" = T)
knitr::opts_chunk$set(
	fig.width=7, fig.height=4, results='asis', echo=F,
	fig.align="center", message=F
)
```

```{r load}
pkgs <- c("magrittr", "kableExtra", "ggpubr")
pkgs.load <- sapply(pkgs, library, character.only=TRUE)
load("01-analysis.RData")
```

# Introduction

- What do we know about **COVID-19**?
- How do we **solve** current issue?
- Potentially arising **problem**?

## COVID-19?

- Current Pandemic
- Widely affecting global population
- Highly transmissible

## Solution?

- **Initial** public health measure: individual vs community-level
- Drug and vaccine development
- Convalescent plasma

## Further Problem? {.build}

* Immediate psychological impact: wariness and negative percepts
- Community quarantine: induces distress
- Socioeconomic instability during quarantine

### Aim

Quantify psychological distress among residents

### Objectives

- Measure **DASS-21** as an initial assessment of psychological distress
- Collect socio-demographic data

# Method

- Study design
- Addressing variables
- Data analysis

## Study Design

- Cross-sectional, analytic-observational
- Targeted respondents: Jakarta, Bogor, Depok, Tangerang, Bekasi
- $\geq$ 18 years of age
- Snowball sampling, stop at 2,000 entries

### Socio-demographic data

- Gender
- Age
- Educational Background

## Addressing Variables {.build .smaller}

### Gender

Assumes binary gender classification

### Age

- Group: Early, prime, mature, elderly
- Based on workforce participation age classification by the UN

### Educational Background

- Low: Elementary, Junior, High School and Vocational degree
- Moderate: Bachelor's, Master's, Professional degree
- High: Doctoral and Post-Doctoral
- Grouping based on proportional convenient segregation

## Data Analysis

- Bivariate: Welch's T-Test and ANOVA
- Generalized Linear Model:
  - Gaussian family, identity link function
  - Response variable underwent square-root transformation
  - VIF: Assess multicollinearity
  - Harrison-McCabe: Determine homogeneity

# Results

- Demographic
- DASS-21 overview
- DASS-21 and associated variables
- Multiple regression

## Demographic

```{r demographic}
edu.group <- df$edu.group %>% table
edu.perc <- edu.group/sum(edu.group)*100
income.group <- df
```

- Total `r dim(df)[1]` participants, `r sum(df$sex=="Male")` men and
  `r sum(df$sex=="Female")` women
- Participants spent `r mean(df$outside)` hours outside
- Age median `r median(df$age)` (18--`r max(df$age)` years old)

### Education

- Low: **`r edu.group[["Low"]]`** (`r edu.perc[["Low"]]`%)
- Moderate: **`r edu.group[["Moderate"]]`** (`r edu.perc[["Moderate"]]`%)
- High: **`r edu.group[["High"]]`** (`r edu.perc[["High"]]`%)

## DASS-21 Overview

Mild distress to none at all; Low DASS-21 score

```{r tbl.dass}
tbl.dass <- data.frame(
	"Depression" = c(
		mean(df$dass$dep > 4)*100,
		mean(df$dass$dep > 6)*100,
		mean(df$dass$dep > 10)*100
	),
	"Anxiety" = c(
		mean(df$dass$anx > 3)*100,
		mean(df$dass$anx > 5)*100,
		mean(df$dass$anx > 8)*100
	),
	"Stress" = c(
		mean(df$dass$str > 7)*100,
		mean(df$dass$str > 9)*100,
		mean(df$dass$str > 12)*100
	)
) %>% set_rownames(c("Mild", "Moderate", "Severe"))

kable(tbl.dass, "html") %>% kable_styling(position="left", font_size=18,
	bootstrap_options="striped", row_label_position="c", full_width=FALSE
) %>% footnote(general="Value as percentage") %>%
	column_spec(2:4, width="8.5em")
```

## DASS-21 and Gender

```{r plt.dass.sex, fig.height=4}

var.list <- paste("dass", names(df$dass), sep="$")
lab.list <- c("Depression", "Anxiety", "Stress", "All")
plt.dass.sex <- function(var.list, lab.list) {
	ggviolin(df, x="sex", y=var.list, color="#458588", fill="#458588",
		xlab="Gender", ylab=lab.list,
		add="mean_sd", add.params=list(color="white")
	) +
	stat_compare_means(size=3, label.x=1.2, method="t.test") +
	font("axis.title", size=9) + font("axis.text", size=9) +
	font("legend.title", size=9) + font("legend.text", size=9)
}
plt.sex1 <- plt.dass.sex(var.list[[1]], lab.list[[1]])
plt.sex2 <- plt.dass.sex(var.list[[2]], lab.list[[2]])
plt.sex3 <- plt.dass.sex(var.list[[3]], lab.list[[3]])
plt.sex4 <- plt.dass.sex(var.list[[4]], lab.list[[4]])
ggarrange(plt.sex1, plt.sex2, plt.sex3, plt.sex4, nrow=2, ncol=2)

```

## DASS-21 and Education

```{r plt.das.edu, fig.height=4}

plt.dass.edu <- function(var.list, lab.list) {
	ggviolin(df, x="edu.group", y=var.list, color="#458588", fill="#458588",
		xlab="Education Level", ylab=lab.list,
		add="mean_sd", add.params=list(color="white")
	) +
	stat_compare_means(size=3, label.x=1.2, method="anova") +
	font("axis.title", size=9) + font("axis.text", size=9) +
	font("legend.title", size=9) + font("legend.text", size=9)
}
plt.edu1 <- plt.dass.edu(var.list[[1]], lab.list[[1]])
plt.edu2 <- plt.dass.edu(var.list[[2]], lab.list[[2]])
plt.edu3 <- plt.dass.edu(var.list[[3]], lab.list[[3]])
plt.edu4 <- plt.dass.edu(var.list[[4]], lab.list[[4]])
ggarrange(plt.edu1, plt.edu2, plt.edu3, plt.edu4, nrow=2, ncol=2)

```

## DASS-21 and Age Group

```{r plt.das.age, fig.height=4}

plt.dass.age <- function(var.list, lab.list) {
	ggviolin(df, x="age.group", y=var.list, color="#458588", fill="#458588",
		xlab="Age Group", ylab=lab.list,
		add="mean_sd", add.params=list(color="white")
	) +
	stat_compare_means(size=3, label.x=2.5, method="anova") +
	font("axis.title", size=9) + font("axis.text", size=9) +
	font("legend.title", size=9) + font("legend.text", size=9)
}
plt.age1 <- plt.dass.age(var.list[[1]], lab.list[[1]])
plt.age2 <- plt.dass.age(var.list[[2]], lab.list[[2]])
plt.age3 <- plt.dass.age(var.list[[3]], lab.list[[3]])
plt.age4 <- plt.dass.age(var.list[[4]], lab.list[[4]])
ggarrange(plt.age1, plt.age2, plt.age3, plt.age4, nrow=2, ncol=2)

```

## Multiple Regression

```{r dass.mv}
lab.tbl.dass.mv <- "\\label{tbl.dass.mv} Summary of Generalized Linear Model
with Maximum Likelihood Estimators"

sum.dass.mv <- dass.mv %>% summary()
aov.dass.mv <- anova(dass.mv, test="F")
tbl.dass.mv <- data.frame(sum.dass.mv$coefficients) %>%
	set_names(colnames(sum.dass.mv$coefficients))

rownames(tbl.dass.mv) %<>% {gsub(
	"sex|\\.group|age|edu",
	"",
	x=.
)} %>% {gsub(":", " -- ", x=.)}

tbl.sig <- tbl.dass.mv %>% subset(subset={.["Pr(>|t|)"] < 0.05})

dass.vif <- car::vif(dass.mv) %>% as.data.frame()
dass.hmc <- lmtest::hmctest(dass.mv)

```

```{r tbl.dass.mv}

dass.fn <- list(
	paste("VIF: {",
		dass.vif$GVIF %>% round(2) %>% paste(collapse=", "),
		"}", sep=""
	),
	paste("Harrison-McCabe, HMC: ", dass.hmc$statistic %>% round(2),
		" p=", dass.hmc$p.value %>% round(2), sep=""
	)
)

kable(tbl.dass.mv, "html") %>%
	kable_styling(full_width=FALSE, font_size=16, position="left") %>%
	pack_rows("Age Group", 3, 5, indent=T) %>%
	pack_rows("Education", 6, 7, indent=T) %>%
	footnote(general=dass.fn, general_title="")
```

# Discussion

- **Sex** dimorphism on distress perception
- **Socio-demographic** status and psychological well-being
- **Age** and quality of life

## Gender and Distress

- We denoted higher psychological distress in women
- Adversities re-experiencing and hyper-arousal
- Negative alterations: cognition and emotion
- Gender-based violence
- Tendency of having longer working period

## Educational Level {.smaller}

- Depressive disorder is less prevalent in well-educated subjects
- Subjects with lower education levels tends to have worse disease
  comprehension
- Education determines depression or anxiety risk
- Lower educational status corresponds to elevated depressive symptoms

### Intervening Variables

- Situational comprehension
- Planning
- Perception
- Information discernment

## Age {.build .smaller}

- Current finding does not follow previous studies
- Suicide rate among elderly rises after SARS pandemic in Hongkong
- COVID-19-infected older adults complaint increased distress

### Rationale

- Different cultural and demographic situation
- An investigation by @Lima2020 only considered infected subjects, while we are evaluating
  healthy subjects
- Referring to Maslow's hierarchy of needs, younger and productive subjects
  perceive more distress
- Erik Erikson: quarantine as burdensome distress

## Limitation

- Non-probability sampling
- Does not fully represent subjects of lower socioeconomic status
- Potential intervening variables

## Conclusion

- Sub-groups at risk of having psychological distress:
  - Women
  - Early to prime age group
  - Low to moderate education level
- Generalizable? Yes, if we only consider VIF and HMC, but...
- Policymaker should consider Maslow's hierarchy of needs in imposing
  quarantine
